---
title: "AclustsCCA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AclustsCCA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

```{r, warning=F, message=F}
devtools::install_github("https://github.com/jennyjyounglee/AclustsCCA")
library("AclustsCCA")
```

## Usage

Below is a list of input parameters:

1. Input data:
**`X``** \eqn{n} by \eqn{p} exposure data matrix, where \eqn{n} is sample size and \eqn{p} is number of exposures.
**`Y``** \eqn{q} by \eqn{n} outcome data matrix, where \eqn{n} is sample size and \eqn{q} is number of outcomes.
**`clusters.list`** A list of clusters with CpG sites obtained using A-clustering, each item is a cluster that contains a set of probes. A-clustering is implemented if `NULL` or can be provided by users.
**`annot`** A preloaded annotation file that includes columns "IlmnID", "Coordinate_37", "Islands_Name", "Relation_to_Island", "UCSC_RefGene_Name". Only needed if \cite{clusters.list} is `NULL`.


2. Input parameters associated with A-clustering:
**`dist.type`** A type of similarity distance function. Options are "spearman" (default), "pearson" (correlation measures) or "euclid". 
**`Aclust.method`** A type of clustering function. Options are "single", "complete" or "average" (default). 
**`thresh.dist`** A similarity distance threshold. Two neighboring clusters are merged to a single cluster if the similarity distance between them is above dist.thresh. Corresponds to $\bar{D}$ in the paper and the default is $0.2$
**`max.dist`** Optional maximum length between neighboring variables permitting to cluster them together. Corresponds to $\bar{d}_{bp}$ in the paper and the default is $1000$.
**`bp.thresh.dist`** A distance in chromosomal location. Any set of methylation sites within an interval smaller or equal to bp.dist will be potentially merged, depending on the similarity between sites at the ends of the interval. Corresponds to $\underline{d}_{bp}$ in the paper and the default is $999$.

3. Input parameters associated with SparseCCA:
**`Xmethod`** A penalty function for the exposure, i.e. penalty function when regressing Y onto X. Options are "lasso", "alasso","gglasso", and "SGL" (default).
**`Ymethod`** A penalty function for the outcome, i.e. penalty function when regressing X onto Y. Options are "lasso", "alasso","gglasso", "SGL", and "OLS" (default).
**`init.method`**         : Initialization method. Options are "lasso", "OLS", and "SVD" (default).
**`X.groupidx`**          : A vector of length \eqn{p} that indicates grouping structure of exposure \eqn{X}.
**`Y.groupidx`**          : A vector of length \eqn{q} that indicates grouping structure of outcome \eqn{Y}.
**`standardize `**        : A logical flag for exposure \eqn{X} and outcome \eqn{Y} standardization, prior to fitting the model.
**`max.iter`**            : A maximum number of iterations of SparseCCA. The default is $100$.
**`conv`**                : A tolerance value for convergence \eqn{epsilon} of SparseCCA. The default is $10e-2$.

4. Input parameters associated permutation test for AclustsCCA:
**`maxnum`** A maximal total number of permutations across all the clusters.
**`maxB`** A maximal number of permutations for a single cluster.
**`permute.tmp.filepath`** Filepath to save intermittent permutation results.
**`permute`** A logical flag for whether to run permutation test or not.
**`nthread`** A number of threads to parallelize permutation test and implementation of SparseCCA across all the clusters.
**`FDR.thresh`** FDR threshold. The default is $0.05$.

## How to implement AclustsCCA

```{r}
data(annot) # row: CpG sites
data.list <- generate.data(n=500)

DATA.X <- data.list$DATA.X # row: subjects (n), column: Metals (p)
DATA.Y <- data.list$DATA.Y # row: subjects (n), column: CpG sites (q)


dist.type <- "spearman"
Aclust.method <-"average"
thresh.dist <-0.2
max.dist <-1000
bp.thresh.dist <-999

Xmethod <- "lasso"
Ymethod <- "OLS"
maxB <- 300

AclustsCCA.result <- AclustsCCA(X=DATA.X,
                                Y=DATA.Y,
                                clusters.list=NULL,
                                annot=annot,
                                # parameters for A-clustering
                                dist.type = dist.type,
                                Aclust.method = Aclust.method,
                                thresh.dist = thresh.dist,
                                max.dist = max.dist,
                                bp.thresh.dist = bp.thresh.dist,
                                # parameters for SparseCCA
                                Xmethod=Xmethod,
                                Ymethod=Ymethod,
                                # parameters for permutation test for AclustsCCA
                                maxB=maxB,
                                permute=F,
                                nthread=3)                         
```

