---
title: "Possible Errors"
author: "Jenny Lee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_to_errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## 1. Common errors one may encounter..

This document summarizes possible error one may encounter when using **`AclustsCCA`**

```{r, eval=F}
# Load annotation file
data(annot) # row: CpG sites
# Load sample data
data(sample.data)
DATA.X <- sample.data$DATA.X # row: subjects (n), column: exposures (p)
DATA.Y <- sample.data$DATA.Y # row: subjects (n), column: CpG sites (q)

# Settings for Aclust
dist.type <- "spearman"
Aclust.method <- "average"
dist.thresh <- 0.2
bp.thresh.clust <- 1000
bp.merge <- 999
```

## 2. Annotation file

A preloaded annotation file should be \code{data.table} and should include columns "IlmnID", "CHR", "Coordinate_37", "Islands_Name", "Relation_to_Island", "UCSC_RefGene_Name".

```{r,error=TRUE}
data(annot)
class(annot)

# What if annotation file is data.frame? 
annot2 <- data.frame(annot)
class(annot2)
all.clusters.list <- Aclust::assign.to.clusters(betas = t(DATA.Y),
                                                annot = annot2,
                                                dist.type = dist.type,
                                                method = Aclust.method,
                                                dist.thresh = dist.thresh,
                                                bp.thresh.clust = bp.thresh.clust,
                                                bp.merge = bp.merge)
```

An annotation file should be ordered by chromosome and chromosomal coordinates of the CpG.
```{r,error=TRUE}
# What if annotation file is not ordered?
annot2 <- annot[sample(1:nrow(annot)),]
all.clusters.list <- Aclust::assign.to.clusters(betas = t(DATA.Y),
                                                annot = annot2,
                                                dist.type = dist.type,
                                                method = Aclust.method,
                                                dist.thresh = dist.thresh,
                                                bp.thresh.clust = bp.thresh.clust,
                                                bp.merge = bp.merge)
str(all.clusters.list)
```
