---
title: "Step 2: Obtain Partial Residuals"
author: "Jenny Lee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_to_partial_residuals}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## 1. What is Partial Residuals?

Partial residuals are .... [@sofer2012multivariate]

## 2. Input Parameters for Partial Residuals

| **`data`** A data matrix with \eqn{n} rows.

| **`Z`** A \eqn{n} by \eqn{r} confounder data matrix, where \eqn{n} is sample size and \eqn{r} is number of potential confounders.

## 3. Obtain partial residuals to adjust for potential confounders

When implementing SparseCCA, partial residuals are used to adjust for potential confounders. Again, computing partial residuals have to be ran only once. I personally suggest to run this part separately and save it for future use.

```{r,eval=F}
# Load sample data
data(sample.data)
DATA.X <- sample.data$DATA.X # row: subjects (n), column: exposures (p)
DATA.Y <- sample.data$DATA.Y # row: subjects (n), column: CpG sites (q)
DATA.Z <- sample.data$DATA.Z # row: subjects (n), column: confounders (r)

X.resid <- partial.residual(data=DATA.X,Z=DATA.Z,nthread=1)
Y.resid <- partial.residual(data=DATA.Y,Z=DATA.Z,nthread=1)
save(X.resid,Y.resid,"partial.residual.RData")
```

## References
