---
title: "Step 3: Implement SparseCCA on each clusterg"
author: "Jenny Lee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_to_sparsecca}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## 1. What is SparseCCA?

SparseCCA is a statistical method .... [@wilms2016robust, @mai2019iterative].

Fast permutation test ... [@gandy2014mmctest, @gandy2019r]

## 2. Input Parameters for SparseCCA

| **`Xmethod`** A penalty function for the exposure, i.e. penalty function when regressing Y onto X. Options are "lasso", "alasso","gglasso", and "SGL" (default).

| **`Ymethod`** A penalty function for the outcome, i.e. penalty function when regressing X onto Y. Options are "lasso", "alasso","gglasso", "SGL", and "OLS" (default).

| **`init.method`**         Initialization method. Options are "lasso", "OLS", and "SVD" (default).

| **`X.groupidx`**          A vector of length \eqn{p} that indicates grouping structure of exposure \eqn{X}.

| **`Y.groupidx`**          A vector of length \eqn{q} that indicates grouping structure of outcome \eqn{Y}.

| **`standardize `**        A logical flag for exposure \eqn{X} and outcome \eqn{Y} standardization, prior to fitting the model.

| **`max.iter`**            A maximum number of iterations of SparseCCA. The default is $100$.

| **`conv`**                A tolerance value for convergence \eqn{epsilon} of SparseCCA. The default is $10e-2$.

| **`maxnum`** A maximal total number of permutations across all the clusters.

| **`maxB`** A maximal number of permutations for a single cluster.

| **`permute.tmp.filepath`** Filepath to save intermittent permutation results.

| **`permute`** A logical flag for whether to run permutation test or not.

| **`nthread`** A number of threads to parallelize permutation test and implementation of SparseCCA across all the clusters.

| **`FDR.thresh`** FDR threshold. The default is $0.05$.

## 3. Implement SparseCCA and permutation test on each cluster identified by A-clustering

As list of clusters are provided and partial residuals are computed, `AclustsCCA` will only run SparseCCA on each cluster identified by A-clustering.

```{r,eval=F}
load("clusters.list.RData")
load("partial.residual.RData")

AclustsCCA.result <- AclustsCCA(X=X.resid,
                                Y=Y.resid,
                                clusters.list=clusters.list,
                                # parameters for SparseCCA
                                Xmethod=Xmethod,
                                Ymethod=Ymethod,
                                X.groupidx=X.groupidx,
                                # parameters for permutation test for AclustsCCA
                                maxB=maxB,
                                permute=TRUE,
                                nthread=nthread) 

summary.AclustsCCA(AclustsCCA.result)
```

## References
